---
header-includes:
- \usepackage{amssymb,amsthm,amsmath}
- \usepackage{chemarr}
output: pdf_document
---

<!--
You can delete the header-includes (lines 3-5 above) if you like and also the chunk below since it is loaded in the skeleton.Rmd file.  They are included so that chap2.Rmd will compile by itself when you hit Knit PDF.
-->

```{r include_acstats_2, include = FALSE}
# This chunk ensures that the acstats package is installed and loaded
# This acstats package includes the template files for the thesis and also
# two functions used for labeling and referencing
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")

if(!require(acstats)){
  library(devtools)
  devtools::install_github("Amherst-Statistics/acstats")
  }
library(acstats)
```


# Example {#typeset-equ}

## Exploring the Data
Data came from Stat 495 final project. (use info from project...). 
Needed a sample of 1000...

Importing the data:

```{r, warning= FALSE}
#using data from final stat 495 project
library(readr)
data_subset <- read_csv("CopyOfdata_subset.csv")

set.seed(1)
#getting a sample of 1000 observations
mysample <- data_subset[sample(1:nrow(data_subset), 1000,
   replace=FALSE),]
```

Picking variables to focus on--> expanding conclusions from Stat 495 project

```{r, warning= FALSE}
#only keeping the variables I want to look at
myvars <- c("Latitude_tri", "Longitude_tri", "poor_or_fair_health", "poor_physical_health_days", "physical_inactivity", "adult_obesity")
smallsample <- mysample[myvars]
```

## Applying CLARA
Step 1: finding k

```{r}
#finding k with project data, using Elbow Method
pkgs <- c("factoextra",  "NbClust")
install.packages(pkgs)

library(factoextra)
library(NbClust)
library(ggplot2)

# Elbow method
fviz_nbclust(new, kmeans, method = "wss") +
    geom_vline(xintercept = 4, linetype = 2)+
  labs(subtitle = "Elbow method")
```

Step 2: Run CLARA function

```{r}
new<- na.omit(smallsample)

## run CLARA
clarasamp <- clara(new[1:6], 4)
```

```{r}
## print components of clarax
print(clarasamp)
summary(clarasamp)
```

```{r}
## plot clusters
plot(new, col = clarasamp$cluster)
## plot centers
points(clarasamp$centers, col = 1:2, pch = 8)
```

```{r} 
#plotting clara
factoextra::fviz_cluster(clarasamp)
```

## Evaluation of CLARA



### Model to Predict Cluster

First, had to include a cluster variable in the original data set, using the data provided by the CLARA function. 

```{r}
#adding each data point's cluster #
cluster<- clarasamp$clustering
cluster_data<- cbind(new, cluster)
```

