---
header-includes:
- \usepackage{amssymb,amsthm,amsmath}
- \usepackage{chemarr}
output: pdf_document
---

<!--
You can delete the header-includes (lines 3-5 above) if you like and also the chunk below since it is loaded in the skeleton.Rmd file.  They are included so that chap2.Rmd will compile by itself when you hit Knit PDF.
-->

# Example {#typeset-equ}

```{r, warning= FALSE}
#loading in packages
library(readr)
library(factoextra)
library(NbClust)
library(ggplot2)
library(cluster)
library(GGally)
```

## Exploring the Data
Data came from Stat 495 final project. (use info from project...). 
Needed a sample of 1000...

Importing the data:

```{r, warning= FALSE}
#using data from final stat 495 project
#library(readr)
data_subset <- read_csv("CopyOfdata_subset.csv")

set.seed(1)
#getting a sample of 1000 observations
mysample <- data_subset[sample(1:nrow(data_subset), 1000,
   replace=FALSE),]
```

Picking variables to focus on--> expanding conclusions from Stat 495 project

```{r, warning= FALSE}
#only keeping the variables I want to look at
myvars <- c("Latitude_tri", "Longitude_tri", "poor_or_fair_health", "poor_physical_health_days", "physical_inactivity", "adult_obesity")
smallsample <- mysample[myvars]
```

## Applying CLARA
Step 1: finding k

```{r}
#finding k with project data, using Elbow Method
#pkgs <- c("factoextra",  "NbClust")
#install.packages(pkgs)

#library(factoextra)
#library(NbClust)
#library(ggplot2)
new<- na.omit(smallsample)
# Elbow method
fviz_nbclust(new, kmeans, method = "wss") +
    geom_vline(xintercept = 4, linetype = 2)+
  labs(subtitle = "Elbow method")
```

Step 2: Run CLARA function

```{r}
#new<- na.omit(smallsample)
#library(cluster)
## run CLARA
clarasamp <- clara(new[1:6], 4)
```

```{r}
## print components of clarax
print(clarasamp)
summary(clarasamp)
```

```{r}
## plot clusters
plot(new, col = clarasamp$cluster)
## plot centers
points(clarasamp$centers, col = 1:2, pch = 8)
```

```{r} 
#plotting clara
factoextra::fviz_cluster(clarasamp)
```

## Evaluation of CLARA



### Model to Predict Cluster

First, had to include a cluster variable in the original data set, using the data provided by the CLARA function. 

```{r}
#adding each data point's cluster #
cluster<- clarasamp$clustering
cluster_data<- cbind(new, cluster)
```

```{r, warning= FALSE}
kitchen_sink<- lm(cluster~., data=cluster_data)
summary(kitchen_sink)

#taking out lat and long
vars <- names(cluster_data) %in% c("Latitude_tri", "Longitude_tri")
cluster_data_new <- cluster_data[!vars]

ggpairs(cluster_data_new)
```

```{r}
fun1<- lm(cluster~ poor_physical_health_days, data= cluster_data_new)
summary(fun1)
```


